---
title: '`Quiz 4` Statistical Inference'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  github_document: default
  html_document:
    self_contained: true
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: true
params:
  author: 'Anderson H Uyekita'
  course: 'Statistical Inference'
  course_url: 'https://www.coursera.org/learn/statistical-inference'
  specialization: 'Data Science: Statistics and Machine Learning Specialization'
  specialization_url: 'https://www.coursera.org/specializations/data-science-statistics-machine-learning'
  instructor: 'Brian Caffo'
  course_start: !r lubridate::ymd('2022/07/01', tz = "America/Sao_Paulo")
  course_finish: !r lubridate::ymd('2022/07/05', tz = "America/Sao_Paulo")
  week: ' Week 4'
always_allow_html: true
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)
library(rmarkdown)
library(lubridate)
```

`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: %s", params$author)}`
`r sprintf(fmt = "* &#x1f4da; Specialization: [%s](%s){target='_blank' rel='noopener'}", params$specialization, params$specialization_url)`
`r sprintf(fmt = "* &#x1f4d6; Course: [%s](%s){target='_blank' rel='noopener'}", params$course, params$course_url)`
    `r sprintf(fmt = "* &#x1F9D1;&#x200d;&#x1F3EB; Instructor: %s", params$instructor)`
`r sprintf(fmt = "* &#x1F4C6; %s", params$week)`
    `r sprintf(fmt = "* &#x1F6A6; Start: %s", format(params$course_start, "%A, %d %B %Y"))`
    `r sprintf(fmt = "* &#x1F3C1; Finish: %s", format(params$course_finish, "%A, %d %B %Y"))`
`r if(!knitr::is_html_output()) {"* &#x1F30E; Rpubs: [Interactive Document](https://rpubs.com/AndersonUyekita/quiz-4_statistical-inference)"}else{"* &#x1F4E6; Github Repository: [Static Document](https://github.com/AndersonUyekita/statistical-inference/blob/main/Week%204/quiz-4_statistical-inference.md){target='_blank' rel='noopener'}"}`

--------------------------------------------------------------------------------

## Question 1

A pharmaceutical company is interested in testing a potential blood pressure lowering medication. Their first examination considers only subjects that received the medication at baseline then two weeks later. The data are as follows (SBP in mmHg)

|Subject|Baseline|Week 2|
|:-----:|:------:|:----:|
|1      |  140   |  132 |
|2      |  138   |  135 |
|3      |  150   |  151 |
|4      |  148   |  146 |
|5      |  135   |  130 |

Consider testing the hypothesis that there was a mean reduction in blood pressure? Give the P-value for the associated two sided T test. 

_(Hint, consider that the observations are paired.)_

- [ ] 0.10
- [x] 0.087
- [ ] 0.05
- [ ] 0.043

**Answer**

* Two weeks later = few observations = T-distribution
* We have the same subject measured in baseline and the experiment = paired

```{r,echo=TRUE}
# Calculating the Hypothesis using a T-distribution with a paired.
q1 <- stats::t.test(x = c(140,138,150,148,135),
                    c(132,135,151,146,130),
                    paired = TRUE,
                    alternative = "two.sided")

# P value.
q1$p.value
```

## Question 2

A sample of 9 men yielded a sample average brain volume of 1,100cc and a standard deviation of 30cc. What is the complete set of values of $\mu_0$ that a test of  $H_0: \mu = \mu_0$ would fail to reject the null hypothesis in a two sided 5% Students t-test?

- [ ] 1081 to 1119
- [ ] 1031 to 1169
- [ ] 1080 to 1120
- [x] 1077 to 1123

**Answer**

* sample ($n$) = 9 
    * Due to the n value being small, it must be used as a t-distribution.
* $\mu$ = 1100 cc
* $\sigma$ = 30 cc
* $\alpha$ = 5%

$\mu \pm t_{(n-1,1-\frac{\alpha}{2}) \cdot \frac{sd}{\sqrt(n)}}$


```{r,echo=TRUE}
n <- 9
mu <- 1100
sd <- 30
alpha <- 5/100

# Calculating the quantile to df = 8 and p = 0.975
t_8_975 <- qt(p = (1 - alpha/2), df = 9 - 1)

# Calculating the CI.
mu + c(-1,1) * t_8_975 * sd / sqrt(n)
```

## Question 3

Researchers conducted a blind taste test of Coke versus Pepsi. Each of four people was asked which of two blinded drinks given in random order that they preferred. The data was such that 3 of the 4 people chose Coke. Assuming that this sample is representative, report a P-value for a test of the hypothesis that Coke is preferred to Pepsi using a one sided exact test.

- [ ] 0.62
- [ ] 0.005
- [ ] 0.10
- [x] 0.31

**Answer**

* p_coke = 3/4
* p_pepsi = 1/4
* Coke is preferred to Pepsi, and;
    Probability to 3 or 4 person prefer Coke.
* Using a one sided exact test.

My Hypothesis:

$$H_{0}: p = 0.5$$
$$H_{a}: p > 0.5$$

General formula:

$$P(n) = {4 \choose n} \cdot (1-p)^{(4-n)} \cdot (p)^{n}$$
$$ P(Prefer \ Coke) = P(p \ge 3)$ = $P(3) + P(4)$$
$$P(p \ge 3) = {4 \choose 3} \cdot (1-p)^{(4-3)} \cdot (p)^{3} + {4 \choose 4} \cdot (1-p)^{(4-4)} \cdot (p)^{4}$$
$$P(p \ge 3) = 0.3125$$

```{r,echo=TRUE}
# Calculating the probability.
P_3_4 <- choose(4,3) * (1-1/2)^1 * (1/2)^3 + choose(4,4) * (1-1/2)^0 * (1/2)^4

P_3_4
```


## Question 4

Infection rates at a hospital above 1 infection per 100 person days at risk are believed to be too high and are used as a benchmark. A hospital that had previously been above the benchmark recently had 10 infections over the last 1,787 person days at risk. About what is the one sided P-value for the relevant test of whether the hospital is *below* the standard? 

- [ ] 0.22
- [x] 0.03
- [ ] 0.52
- [ ] 0.11

**Answer**

```{r,echo=TRUE}
ppois(q = 10, lambda = 1/100 * 1787)
```

## Question 5

Suppose that 18 obese subjects were randomized, 9 each, to a new diet pill and a placebo. Subjects’ body mass indices (BMIs) were measured at a baseline and again after having received the treatment or placebo for four weeks. The average difference from follow-up to the baseline (followup - baseline) was −3 kg/m2 for the treated group and 1 kg/m2 for the placebo group. The corresponding standard deviations of the differences was 1.5 kg/m2 for the treatment group and 1.8 kg/m2 for the placebo group. Does the change in BMI appear to differ between the treated and placebo groups?  Assuming normality of the underlying data and a common population variance, give a p-value for a two sided t test.

- [ ] Less than 0.05, but larger than 0.01
- [x] Less than 0.01
- [ ] Less than 0.10 but larger than 0.05
- [ ] Larger than 0.10

**Answer**

$$H_{0}: \mu_{diff,treated} - \mu_{diff,placebo} = 0$$
$$H_{a}: \mu_{diff,treated} - \mu_{diff,placebo} \ne 0$$

$$\underbrace{\mu_{diff,treated} - \mu_{diff,placebo}}_{0} = \bar{treated} - \bar{placebo} \pm t_{(n + n -2, 1-\frac{\alpha}{2})} \cdot S_{p} \cdot (\frac{1}{n} + \frac{1}{n})^{1/2} $$
$$t_{(n + n -2, 1-\frac{\alpha}{2})} \cdot S_{p} \cdot (\frac{1}{n} + \frac{1}{n})^{1/2} = \bar{treated} - \bar{placebo}$$
$$t_{(n + n -2, 1-\frac{\alpha}{2})} = \frac{\bar{treated} - \bar{placebo}}{S_{p} \cdot (\frac{1}{n} + \frac{1}{n})^{1/2}}$$

```{r,echo=TRUE}
# Sample
n <- 9

# Treated
x_treat <- -3 
s_treat <- 1.5

# Placebo
x_placebo <- 1 
s_placebo <- 1.8

# The pooled variance estimator is:
sp <- sqrt(((n - 1) * s_treat^2 + (n - 1) * s_placebo^2)/(n + n - 2))

# Calculating the quantile given: mu_treat - mu_placebo = 0
t <- (x_treat - x_placebo)/(sp * sqrt(1/n + 1/n))

# Given t, its possible to calculate the probability.
q5_pvalue <- pt(t, n + n - 2)

# Formatting
format(q5_pvalue, digits = 2,scientific = FALSE)
```

## Question 6

Brain volumes for 9 men yielded a 90% confidence interval of 1,077 cc to 1,123 cc. Would you reject in a two sided 5% hypothesis test of $H_0 : \mu = 1,078$? 

- [ ] Where does Brian come up with these questions?
- [x] No you wouldn't reject.
- [ ] Yes you would reject.
- [ ] It's impossible to tell.

**Answer**

The 90% CI already includes the $\mu=1078$, so if I increase the CI to 95%, it will still contain the $\mu=1078$. For this reason, I will not reject the $H_0$ statement.

## Question 7

Researchers would like to conduct a study of $100$ healthy adults to detect a four year mean brain volume loss of $.01~mm^3$. Assume that the standard deviation of four year volume loss in this population is $.04~mm^3$. About what would be the power of the study for a $5\%$ one sided test versus a null hypothesis of no volume loss?

- [ ] 0.70
- [ ] 0.60
- [x] 0.80
- [ ] 0.50

**Answer**

* n = 100 (normal)
* one sided test = `lower.tail = FALSE`
* sd = 0.04
* delta = 0.01
* $\alpha = 5\%$

```{r,echo=TRUE}
# Inputs
n <- 100
alpha <- 5/100
sd <- 0.04
mu <- 0.01

# Calculating the z value from alpha equal to 5%
z <- qnorm(p = (1 - alpha))

# Calculating the probability.
pnorm(z * sd/sqrt(n), mean = mu, sd = sd/sqrt(n), lower.tail = FALSE)
```


## Question 8

Researchers would like to conduct a study of $n$ healthy adults to detect a four year mean brain volume loss of $.01~mm^3$. Assume that the standard deviation of four year volume loss in this population is $.04~mm^3$. About what would be the value of $n$ needed for $90\%$ power of type one error rate of $5\%$ one sided test versus a null hypothesis of no volume loss?

- [ ] 120
- [ ] 180
- [ ] 160
- [x] 140

**Answer**

* $\Delta$ = 0.01
* sd = 0.04
* power = 90%
* type one error = 5% = `sig.level`
* type = one sided test

```{r,echo=TRUE}
# Calculating the n value.
q8 <- power.t.test(power = 0.9,
                   delta = 0.01,
                   sd = 0.04,
                   sig.level = 0.05,
                   type = "one.sample",
                   alternative = "one.sided")

# Printing the n
round(q8$n)
```

## Question 9

As you increase the type one error rate, $\alpha$, what happens to power?

- [ ] No, for real, where does Brian come up with these problems?
- [ ] It's impossible to tell given the information in the problem.
- [ ] You will get smaller power.
- [x] You will get larger power.
