---
title: '`Quiz 2` Statistical Inference'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  github_document: default
  html_document:
    self_contained: yes
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: true
params:
  author: 'Anderson H Uyekita'
  course: 'Statistical Inference'
  course_url: 'https://www.coursera.org/learn/statistical-inference'
  specialization: 'Data Science: Statistics and Machine Learning Specialization'
  specialization_url: 'https://www.coursera.org/specializations/data-science-statistics-machine-learning'
  instructor: 'Brian Caffo'
  course_start: !r lubridate::ymd('2022/06/30', tz = "America/Sao_Paulo")
  course_finish: !r lubridate::ymd('2022/07/01', tz = "America/Sao_Paulo")
  week: ' Week 2'
always_allow_html: true
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)
library(rmarkdown)
library(lubridate)
```

`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: %s", params$author)}`
`r sprintf(fmt = "* &#x1f4da; Specialization: [%s](%s){target='_blank' rel='noopener'}", params$specialization, params$specialization_url)`
`r sprintf(fmt = "* &#x1f4d6; Course: [%s](%s){target='_blank' rel='noopener'}", params$course, params$course_url)`
    `r sprintf(fmt = "* &#x1F9D1;&#x200d;&#x1F3EB; Instructor: %s", params$instructor)`
`r sprintf(fmt = "* &#x1F4C6; %s", params$week)`
    `r sprintf(fmt = "* &#x1F6A6; Start: %s", format(params$course_start, "%A, %d %B %Y"))`
    `r sprintf(fmt = "* &#x1F3C1; Finish: %s", format(params$course_finish, "%A, %d %B %Y"))`
`r if(!knitr::is_html_output()) {"* &#x1F30E; Rpubs: [Interactive Document](https://rpubs.com/AndersonUyekita/quiz-2_statistical-inference)"}else{"* &#x1F4E6; Github Repository: [Static Document](https://github.com/AndersonUyekita/statistical-inference/blob/main/Week%202/quiz-2_statistical-inference.md){target='_blank' rel='noopener'}"}`

--------------------------------------------------------------------------------

## Question 1

What is the variance of the distribution of the average an IID (independent and identically distributed) draw of $n$ observations from a population with mean $\mu$ and variance $\sigma^2$.

- [ ] $\sigma / n$
- [ ] $\sigma^2$
- [x] $\frac{\sigma^2}{n}$
- [ ] $\sigma / \sqrt{n}$

**Answer**

$$Var(\bar X) = \frac{\sigma^2}{n}$$
Where:

* $\sigma^2$: population variance
* $n$: number of observation
* $Var(\bar X)$: Variance of the sample mean

## Question 2

Suppose that diastolic blood pressures (DBPs) for men aged 35-44 are normally distributed with a mean of 80 (mm Hg) and a standard deviation of 10. About what is the probability that a random 35-44 year old has a  DBP less than 70?

- [ ] 22%
- [x] 16%
- [ ] 8%
- [ ] 32%

**Answer**

> normally distributed

```{r}
# option 1
pnorm(q = 70, mean = 80, sd = 10, lower.tail = TRUE)

# option 2
pnorm(q = (70-80)/10, lower.tail = TRUE)
```

## Question 3

Brain volume for adult women is normally distributed with a mean of about 1,100 cc for women with a standard deviation of 75 cc. What brain volume represents the 95th percentile?

- [ ] approximately 977
- [ ] approximately 1175
- [ ] approximately 1247
- [x] approximately 1223

**Answer**

> normally distributed

```{r}
qnorm(p = 0.95, mean = 1100, sd = 75)
```


## Question 4

Refer to the previous question. Brain volume for adult women is about 1,100 cc for women with a standard deviation of 75 cc. Consider the sample mean of 100 random adult women from this population. What is the 95th percentile of the distribution of that sample mean? 

- [x] approximately 1112 cc
- [ ] approximately 1115 cc
- [ ] approximately 1110 cc
- [ ] approximately 1088 cc

**Answer**

> normally distributed

```{r}
qnorm(p = 0.95, mean = 1100, sd = 75/sqrt(100))
```

## Question 5

You flip a fair coin 5 times, about what's the probability of getting 4 or 5 heads?

- [ ] 12%
- [x] 19%
- [ ] 6%
- [ ] 3%

**Answer**

> Binomial Distribution

4 out of 5 coin tosses should be heads:

$$p1 = {5 \choose 4} \cdot (1-p)^{5-4} \cdot (p)^4$$

5 out of 5 coin tosses should be heads:

$$p2 = {5 \choose 5} \cdot (1-p)^{5-5} \cdot (p)^5$$

Thus:

$$R = p1 + p2$$

```{r}
p <- 0.5
n <- 5

P1 <- (factorial(5) / (factorial(5-4) * factorial(4))) * (1 - p)^(n-4) * (p)^(4)
P2 <- (factorial(5) / (factorial(5-5) * factorial(5))) * (1 - p)^(n-5) * (p)^(5)

R <- P1 + P2

R
```

## Question 6

The respiratory disturbance index (RDI), a measure of sleep disturbance, for a specific population has a mean of 15 (sleep events per hour) and a standard deviation of 10. They are not normally distributed. Give your best estimate of the probability that a sample mean RDI of 100 people is between 14 and 16 events per hour?

- [ ] 47.5%
- [ ] 34%
- [ ] 95%
- [x] 68%

**Answer**

> **not** normally distributed

```{r}
p1 <- pnorm((14-15)/(10/sqrt(100)), lower.tail = TRUE)

p2 <- pnorm((16-15)/(10/sqrt(100)), lower.tail = TRUE)

p2 - p1
```

## Question 7

Consider a standard uniform density. The mean for this density is .5 and the variance is $1/12$. You sample 1,000 observations from this distribution and take the sample mean, what value would you expect it to be near?

- [ ] 0.25
- [ ] 0.10
- [ ] 0.75
- [x] 0.5

**Answer**

> Law of large numbers in action (LLN)

$$Var(X) = \sigma^2 = \frac{1}{12} = \frac{(b-a)^2}{12}$$
$$mean = 0.5 =\frac{a+b}{2} $$
$$\begin{cases}
b+a = 1 \\
b-a = 1
\end{cases}$$

$b = 1$ and $a = 0$


```{r}
library(ggplot2)

set.seed(2022)

n <- c(10,50,75,100,300,500,1000,2000,5000,10000, 20000)
df <- data.frame("mean" = rep(NA, length(n)),"n" = rep(NA, length(n)), "var" = rep(NA, length(n)))

for (i in 1:length(n)) {
     df$mean[i] <- mean(runif(n = n[i], min = 0, max = 1))
     df$var[i] <- sd(runif(n = n[i], min = 0, max = 1))^2
     df$n[i] <- n[i]
}

ggplot(data = df, aes(x = n, y = mean)) +
    geom_line() +
    scale_y_continuous(limits = c(0.375, 0.625), breaks = seq(0.25, 0.75, by = 0.125))
```

## Question 8

The number of people showing up at a bus stop is assumed to be Poisson with a mean of 5 people per hour. You watch the bus stop for 3 hours. About what's the probability of viewing 10 or fewer people?

- [x] 0.12
- [ ] 0.03
- [ ] 0.08
- [ ] 0.06

**Answer**

> Poisson Distribution

```{r}
ppois(q = 10,lambda = 5 * 3)
```